Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2018-02-05 aggregation/agg-intro.tex
Streamlined Intro

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.  In the scope of Environmental Product Declarations (EPDs), studies within an industry are supposed to be rendered comparable by adhering to a common product category rule (PCR) \citep{Fet_2006}.  However, this prescription is insufficient to ensure the comparability of different declarations, even if they use the same PCR \citep{Modahl_2012}.  In addition, PCRs are often themselves not unique and can include wide variations in scope and boundary requirements \citep{Subramanian_2012}, leading to ambiguity.

An LCA study result is ultimately an assertion that for some constructed PSM, the delivery of a particular reference flow is associated with a certain amount of environmental impact or potential impact.  But if the PSM itself is ambiguously stated, then the significance of the result is ambiguous as well.  For a computational model to have scientific significance, its results need to be both verifiable and replicable by other scientists \citep{Fomel_2009, Mesirov_2010}.  But in the case of LCA, the structure and contents of the model are never stated precisely; instead they are \emph{described} in a written report, and the description must be interpreted by a reader.  When a study is reviewed, it is often the written report and \emph{not} the PSM itself that is the object of review.  \cite{iso14071}, which provides guidelines on critical review, emphasizes that review of the PSM or individual datasets is optional.

%A common theme in all these controversies is the inability for authors and their readers to formally agree on the structure and content of the PSM.
Although the general structure for life cycle assessment can be stated formally, there is no current practice for formally describing the PSM and thus, no automatic mechanisms for interpretation or review of published LCA studies.  Moreover, there is no current practice that permits a reviewer to verify the correctness of the computations inherent in an LCA study if access to the PSM itself is not available.  The purpose of this paper is to advance a framework for the formal description of product system models that would provide robust transparency and reproducibility of computed results.

>>>>

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.

An LCA study result is ultimately an assertion that for some constructed PSM, the delivery of a particular reference flow is associated with a certain amount of environmental impact or potential impact.  But if the PSM itself is ambiguously stated, then the significance of the result is ambiguous as well.  Often the structure and contents of the model are never stated precisely; instead they are \emph{described} in a written report, and the description must be interpreted by a reader.  When a study is reviewed, it is often the written report and \emph{not} the PSM itself that is the object of review.  \cite{iso14071}, which provides guidelines on critical review, emphasizes that review of the PSM or individual datasets is optional.  
%A common theme in all these controversies is the inability for authors and their readers to formally agree on the structure and content of the PSM.
Although the general structure for life cycle assessment can be stated formally, there is no current practice for formally describing the PSM and thus, no automatic mechanisms for interpretation or review of published LCA studies.  Moreover, there is no current practice that permits a reviewer to verify the correctness of the computations inherent in an LCA study if access to the PSM itself is not available.

Because of the diversity of software and techniques for performing LCA, it is important to develop methods for communicating the structure of the PSM that is not dependent on any particular software or database \cite{Kuczenski_JLCA_2018}.    The purpose of this paper is to advance a framework for the formal description of PSM that would allow for sharing, revision, and critical review of models while also enabling authors to protect the privacy of confidential information.  This objective is also in line with the broader imperative throughout industrial ecology to improve transparency and reproducibility of computed results \cite{Hertwich_JIE_2018}.

### 2018-02-05 aggregation/agg-multioutput.tex
Streamlined Intro

\subsection{Multi-Output Foreground Configurations}

@@

\begin{table}
  \begin{center}
  \caption{Chlor-Alkali Electrolysis, membrane cell [RER] (Ecoinvent 3.2; excerpt)}
  \label{tbl:chlor-alkali}
  \begin{tabular}{rccl}
    \toprule
    \bf Flow & \bf Unit & \bf Direction & \bf Value \\
    \midrule
    \multicolumn{4}{l}{Reference:} \\
    Chlorine, gaseous & kg & Output & 1 \\
    Hydrogen, liquid & kg & Output & 0.028 \\
    Sodium hydroxide & kg dry & Output & 1.13 \\
    \midrule
    \multicolumn{4}{l}{Requirements:} \\
    Electricity, medium voltage & kWh & Input & 2.97 \\
    Sodium chloride, powder & kg & Input & 1.75 \\
    Chemical factory, organics & unit & Input & 4e-10 \\
    Sludge, NaCl electrolysis & kg & Output & 0.0153 \\[1ex]
    \multicolumn{4}{l}{Emissions:} \\
    Chloride [16887-00-6], surface water & kg & Output & 0.015 \\
    Carbon dioxide [124-38-9], urban air & kg & Output & 0.0031 \\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}
  \begin{center}
  \caption{Chlor-Alkali Electrolysis as a foreground model, partitioning allocation}
  \label{tbl:partition}
  \footnotesize\sffamily
  \begin{tabular}{l|ccc}
    \midrule
    \bf $A_f$ & 0 & 1 & 2 \\
    \midrule
    \textbf{Chlorine, gaseous (kg)} & $\ast$ & & \\
    Hydrogen, liquid (kg) & -0.028 & $\ast$ & \\
    Sodium hydroxide (kg dry) & -1.13 & & $\ast$ \\
    \midrule
    \bf $A_d$ & 0 & 1 & 2 \\
    \midrule
    Electricity, medium voltage (kWh) & 2.97 & 1.38 & 1.38 \\
    Sodium chloride, powder (kg) & 1.75 & 0.811 & 0.811 \\
    Chemical factory, organics (unit) & 4e-10 & 1.9e-10 & 1.9e-10 \\
    Sludge, NaCl electrolysis (kg) & 0.0153 & 0.00709 & 0.00709 \\
    \midrule
    \bf $B_f$ & 0 & 1 & 2 \\
    \midrule
    Chloride, to surface water (kg) & 0.015 & 0.00695 & 0.00695 \\
    Carbon dioxide, to urban air (kg) & 0.0031 & 0.00144 & 0.00144 \\
    \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}[h]
  \begin{center}
  \caption{Chlor-Alkali Electrolysis as a foreground model, classical system expansion}
  \label{tbl:cse}
  \footnotesize\sffamily
  \begin{tabular}{l|ccc}
    \midrule
    \bf $A_f$ & 0 & 1 & 2 \\
    \midrule
    \textbf{Chlorine, gaseous (kg)} & $\ast$ & & \\
    Hydrogen, liquid (kg) & -0.028 & $\ast$ & \\
    Sodium hydroxide (kg dry) & -1.13 & & $\ast$ \\
    \midrule
    \bf $A_d$ & 0 & 1 & 2 \\
    \midrule
    Electricity, medium voltage (kWh) & 2.97 &  &  \\
    Sodium chloride, powder (kg) & 1.75 &  & \\
    Chemical factory, organics (unit) & 4e-10 &  &  \\
    Sludge, NaCl electrolysis (kg) & 0.0153 &  &  \\
    \midrule
    \bf $B_f$ & 0 & 1 & 2 \\
    \midrule
    Chloride, to surface water (kg) & 0.015 &  &  \\
    Carbon dioxide, to urban air (kg) & 0.0031 &  &  \\
    \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}[t]
  \begin{center}
  \caption{Chlor-Alkali Electrolysis as a foreground model, consequential system expansion.  An inventory for the electrolysis of water (shown in blue) is provided as an alternate activity to produce hydrogen, while the sodium hydroxide product is assumed to substitute for for magnesium hydroxide (shown in red).}
  \label{tbl:exp}
  \footnotesize\sffamily
  \psset{unit=0.5cm}
  \begin{tabular}{l|cccc}
    \midrule
    \bf $A_f$ & 0 & 1 & 2 & \blue\textit{3} \\
    \midrule
    \textbf{Chlorine, gaseous (kg)} & $\ast$ & & &\\
    Hydrogen, liquid (kg) & -0.028 & $\ast$ & &\\
    Sodium hydroxide (kg dry) & -1.13 & & $\ast$ & \\
    \blue\textit{Hydrogen, electrolysis (kg)} & & \blue\textit{1} & & $\ast$ \\
    \midrule
    \bf $A_d$ & 0 & 1 & 2 & \blue\textit{3}\\
    \midrule
    Electricity, medium voltage (kWh) & 2.97 &  & & \blue\textit{50} \\
    Sodium chloride, powder (kg) & 1.75 &  & & \\
    Chemical factory, organics (unit) & 4e-10 &  & &  \\
    Sludge, NaCl electrolysis (kg) & 0.0153 &  & & \\
    \blue\textit{Water, distilled} (kg) & & & & \blue\textit{9} \\
    \red\textit{Magnesium hydroxide (kg dry)} & & & \red\textit{0.8} & \\
    \midrule
    \bf $B_f$ & 0 & 1 & 2 & \blue\textit{3}\\
    \midrule
    Chloride, to surface water (kg) & 0.015 &  & & \\
    Carbon dioxide, to urban air (kg) & 0.0031 &  & & \\
    \midrule
  \end{tabular}
  \end{center}
\end{table}

>>>>

\subsection{Multi-Output Foreground Processes}

@@

\begin{table}
  \begin{center}
  \caption{Example: Chlor-alkali electrolysis, membrane cell [RER] (Ecoinvent 3.2; excerpt)}
  \label{tbl:chlor-alkali}
  \begin{tabular}{rccl}
    \toprule
    \bf Flow & \bf Unit & \bf Direction & \bf Value \\
    \midrule
    \multicolumn{4}{l}{Reference:} \\
    Chlorine, gaseous & kg & Output & 1 \\
    Hydrogen, liquid & kg & Output & 0.028 \\
    Sodium hydroxide & kg dry & Output & 1.13 \\
    \midrule
    \multicolumn{4}{l}{Requirements:} \\
    Electricity, medium voltage & kWh & Input & 2.97 \\
    Sodium chloride, powder & kg & Input & 1.75 \\
    Chemical factory, organics & unit & Input & 4e-10 \\
    Sludge, NaCl electrolysis & kg & Output & 0.0153 \\[1ex]
    \multicolumn{4}{l}{Emissions:} \\
    Chloride [16887-00-6], surface water & kg & Output & 0.015 \\
    Carbon dioxide [124-38-9], urban air & kg & Output & 0.0031 \\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}
  \begin{center}
  \caption{Chlor-alkali electrolysis as a foreground model, partitioning allocation}
  \label{tbl:partition}
  \footnotesize\sffamily
  \begin{tabular}{l|ccc}
    \midrule
    \bf $A_f$ & 0 & 1 & 2 \\
    \midrule
    \textbf{Chlorine, gaseous (kg)} & $\ast$ & & \\
    Hydrogen, liquid (kg) & -0.028 & $\ast$ & \\
    Sodium hydroxide (kg dry) & -1.13 & & $\ast$ \\
    \midrule
    \bf $A_d$ & 0 & 1 & 2 \\
    \midrule
    Electricity, medium voltage (kWh) & 2.97 & 1.38 & 1.38 \\
    Sodium chloride, powder (kg) & 1.75 & 0.811 & 0.811 \\
    Chemical factory, organics (unit) & 4e-10 & 1.9e-10 & 1.9e-10 \\
    Sludge, NaCl electrolysis (kg) & 0.0153 & 0.00709 & 0.00709 \\
    \midrule
    \bf $B_f$ & 0 & 1 & 2 \\
    \midrule
    Chloride, to surface water (kg) & 0.015 & 0.00695 & 0.00695 \\
    Carbon dioxide, to urban air (kg) & 0.0031 & 0.00144 & 0.00144 \\
    \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}[h]
  \begin{center}
  \caption{Chlor-alkali electrolysis as a foreground model, classical system expansion}
  \label{tbl:cse}
  \footnotesize\sffamily
  \begin{tabular}{l|ccc}
    \midrule
    \bf $A_f$ & 0 & 1 & 2 \\
    \midrule
    \textbf{Chlorine, gaseous (kg)} & $\ast$ & & \\
    Hydrogen, liquid (kg) & -0.028 & $\ast$ & \\
    Sodium hydroxide (kg dry) & -1.13 & & $\ast$ \\
    \midrule
    \bf $A_d$ & 0 & 1 & 2 \\
    \midrule
    Electricity, medium voltage (kWh) & 2.97 &  &  \\
    Sodium chloride, powder (kg) & 1.75 &  & \\
    Chemical factory, organics (unit) & 4e-10 &  &  \\
    Sludge, NaCl electrolysis (kg) & 0.0153 &  &  \\
    \midrule
    \bf $B_f$ & 0 & 1 & 2 \\
    \midrule
    Chloride, to surface water (kg) & 0.015 &  &  \\
    Carbon dioxide, to urban air (kg) & 0.0031 &  &  \\
    \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}[t]
  \begin{center}
  \caption{Chlor-alkali electrolysis as a foreground model, consequential system expansion.  An inventory for the electrolysis of water (shown in blue) is provided as an alternate activity to produce hydrogen, while the sodium hydroxide product is assumed to substitute for for magnesium hydroxide (shown in red).}
  \label{tbl:exp}
  \footnotesize\sffamily
  \psset{unit=0.5cm}
  \begin{tabular}{l|cccc}
    \midrule
    \bf $A_f$ & 0 & 1 & 2 & \blue\textit{3} \\
    \midrule
    \textbf{Chlorine, gaseous (kg)} & $\ast$ & & &\\
    Hydrogen, liquid (kg) & -0.028 & $\ast$ & &\\
    Sodium hydroxide (kg dry) & -1.13 & & $\ast$ & \\
    \blue\textit{Hydrogen, electrolysis (kg)} & & \blue\textit{1} & & $\ast$ \\
    \midrule
    \bf $A_d$ & 0 & 1 & 2 & \blue\textit{3}\\
    \midrule
    Electricity, medium voltage (kWh) & 2.97 &  & & \blue\textit{50} \\
    Sodium chloride, powder (kg) & 1.75 &  & & \\
    Chemical factory, organics (unit) & 4e-10 &  & &  \\
    Sludge, NaCl electrolysis (kg) & 0.0153 &  & & \\
    \blue\textit{Water, distilled} (kg) & & & & \blue\textit{9} \\
    \red\textit{Magnesium hydroxide (kg dry)} & & & \red\textit{0.8} & \\
    \midrule
    \bf $B_f$ & 0 & 1 & 2 & \blue\textit{3}\\
    \midrule
    Chloride, to surface water (kg) & 0.015 &  & & \\
    Carbon dioxide, to urban air (kg) & 0.0031 &  & & \\
    \midrule
  \end{tabular}
  \end{center}
\end{table}
